---
title: "Historic data in vehicle + AZD4625-treated P53 GEMMs"
subtitle: Statistical analysis of all data
author: "Gabriel Teku"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(emmeans)
library(rstan)
library(rstanarm)
library(bayesplot)
library(brms)
library(loo)

set.seed <- 1234

## colourblind-friendly colours for ggplot
cbb_palette <- c(
  "#000000", "#E69F00", "#56B4E9", "#F0E442", "#0072B2",
  "#D55E00", "#CC79A7", "#0099C6", "#C7A92A", "#009E73",
  "#5A9F9F", "#F8766D", "#C49A00", "#F2C80F", "#993D3D",
  "#4F5D75", "#0089A3", "#A8A878", "#63C5DA", "#79A779"
)

options(digits = 5, tibble.print_max = 10^6)

if(file.exists("03-data.RData")) load("03-data.RData")
```


# Load cleaned data

```{r load-cleaned-data, echo=FALSE}
dt <- read_rds("clean-data.rds")
dt <- dt %>% 
    mutate(mutation = relevel(mutation, ref = "p53"),
           trt = relevel(trt, ref = "Vehicle"),
           timepoint = relevel(timepoint, "baseline")
    )
```


# Statistical inference 

## Fit Bayesian mixed effects model 

Interested in picking up trt effects & effects of treatment over time. 
Does this differ between mutation statuses?
Consider study as random effect with sampling from the entire study space.

### Vehicle vs AZD4625 mono

```{r fit-vehicle-vs-azd4625mono}
# Extract subset of data
mod_df_4625mono <- dt %>% 
    filter(trt %in% c("Vehicle", "AZD4625")) %>% 
    # filter(!timepoint == "3weeks") %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove)) %>% 
    droplevels()
studies_df_4625mono <- mod_df_4625mono %>% distinct(study, mutation)
studies_df_4625mono

mod_df_4625mono_t1_t2 <- mod_df_4625mono %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625mono_t1_t2 <- mod_df_4625mono_t1_t2 %>% distinct(study, mutation)
studies_df_4625mono_t1_t2

# Fit model
mod_4625mono <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                          data = mod_df_4625mono,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 5000)
mod_4625mono_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                          data = mod_df_4625mono_t1_t2,
                          adapt_delta = 0.999,
                          seed = 1234,
                          iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625mono)
summary(mod_4625mono_t1_t2)

## Posterior Predictive Checks
pp_check(mod_4625mono)
pp_check(mod_4625mono_t1_t2)

## Generate trace plots
posterior_mod_4625mono <- as.matrix(mod_4625mono)
mcmc_trace(posterior_mod_4625mono, regex_pars = "trt")

posterior_mod_4625mono_t1_t2 <- as.matrix(mod_4625mono_t1_t2)
mcmc_trace(posterior_mod_4625mono_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_4625mono)
stan_diag(mod_4625mono_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625mono_ar <- mcmc_areas_ridges(posterior_mod_4625mono, 
                       regex_pars = "trt",
                       prob = 0.95) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625 mono vs Vehicle: all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625mono_ar
ggsave("figures/fit-posterior-azd4625-vehicle.png", plt_4625mono_ar)

plt_4625mono_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_4625mono, 
    pars = vars(starts_with("trt"),
                -c("trtAZD4625:timepoint2weeks:mutationstk11")
    ),
    prob = 0.95
    ) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625 mono vs Vehicle: all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold", size = 18)) 
plt_4625mono_ar_no3way
ggsave("figures/fit-posterior-azd4625-vehicle-no3way.png", 
       plt_4625mono_ar_no3way)
## plot baseline and 2 weeks
plt_4625mono_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625mono_t1_t2,
                                           regex_pars = "trt",
                                           prob = 0.95) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625 mono vs Vehicle: baseline and 2 weeks, all studies") +
    ggtitle("AZD4625 mono vs Vehicle: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold", size = 15))
plt_4625mono_ar_t1_t2
ggsave("figures/fit-posterior-azd4625mono-vehicle.png", plt_4625mono_ar_t1_t2)

## plot baseline and 2 weeks without 3-way interactions
plt_4625mono_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_4625mono_t1_t2,
    pars = vars(starts_with("trt"),
                - "trtAZD4625:timepoint2weeks:mutationstk11"
    ),
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625 mono vs Vehicle: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    ggtitle("AZD4625 mono vs Vehicle: baseline and 2 weeks") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625mono_ar_t1_t2_no3way
ggsave("figures/fit-posterior-azd4625mono-vehicle.png", 
       plt_4625mono_ar_t1_t2_no3way)
save.image("03-data.RData")
```



### Vehicle vs AZD4625 combo

```{r fit-model-vehicle-vs-4625combo}
# Extract subset of data
mod_df_4625combo <- dt %>% 
    filter(str_detect(trt, c("Vehicle|AZD4625 +"))) %>%
     # filter(!timepoint == "3weeks") %>%  
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625combo <- mod_df_4625combo %>% distinct(study, mutation)
#
mod_df_4625combo_t1_t2 <- mod_df_4625combo %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625combo_t1_t2 <- mod_df_4625combo_t1_t2 %>% 
    distinct(study, mutation)

# Fit model
mod_4625combo <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                               (1|mutation/study),
                           data = mod_df_4625combo,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 6000)
mod_4625combo_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                          data = mod_df_4625combo_t1_t2,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625combo) %>% head()
summary(mod_4625combo) %>% tail()
summary(mod_4625combo_t1_t2) %>% head()
summary(mod_4625combo_t1_t2) %>% tail()
mod_4625combo
mod_4625combo_t1_t2

## Posterior Predictive Checks
pp_check(mod_4625combo)
pp_check(mod_4625combo_t1_t2)

## Model performance
# (loo_4625combo <- loo(mod_4625combo, k_threshold = 0.7))
(loo_4625combo_t1_t2 <- loo(mod_4625combo_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_4625combo <- as.matrix(mod_4625combo)
mcmc_trace(posterior_mod_4625combo, regex_pars = "trt")

posterior_mod_4625combo_t1_t2 <- as.matrix(mod_4625combo_t1_t2)
mcmc_trace(posterior_mod_4625combo_t1_t2, regex_pars = "trt")

## Model performance
(loo_4625combo_t1_t2 <- loo(mod_4625combo_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625combo)
stan_diag(mod_4625combo_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625combo_ar <- mcmc_areas_ridges(posterior_mod_4625combo, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("AZD4625 combo vs Vehicle: all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625combo_ar 
ggsave("figures/fit-posterior-azd4625combo-vs-vehicle.png", plt_4625combo_ar)

plt_4625combo_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_4625combo, 
    pars = vars(starts_with("trt"),
                -"trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11"
    ), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625 combo vs Vehicle: all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625combo_ar_no3way 
ggsave("figures/fit-posterior-azd4625combo-vs-vehicle-no3way.png",
       plt_4625combo_ar_no3way)

plt_4625combo_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625combo_t1_t2, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("AZD4625 combo vs Vehicle: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625combo_ar_t1_t2 
ggsave("figures/fit-posterior-azd4625combo-vs-vehicle-t1-t2.png",
       plt_4625combo_ar_t1_t2)

plt_4625combo_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_4625combo_t1_t2, 
    pars = vars(starts_with("trt"),
                -"trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11"
    ),
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625 combo vs Vehicle: baseline and 2 weeks, all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625combo_ar_t1_t2_no3way 
ggsave("figures/fit-posterior-azd4625combo-vs-vehicle-t1-t2-no3way.png",
       plt_4625combo_ar_t1_t2_no3way)

save.image("03-data.RData")
```



### Vehicle vs AZD4625 mono + combo

```{r fit-model-vehicle-vs-4625moco}
# Extract subset of data
mod_df_4625moco <- dt %>% 
    filter(str_detect(trt, c("Vehicle|AZD4625"))) %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625moco <- mod_df_4625moco %>% distinct(study, mutation)
#
mod_df_4625moco_t1_t2 <- mod_df_4625moco %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625moco_t1_t2 <- mod_df_4625moco_t1_t2 %>% distinct(study, mutation)

# Fit model
mod_4625moco <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                           data = mod_df_4625moco,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)
mod_4625moco_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                                    (1|mutation/study),
                          data = mod_df_4625moco_t1_t2,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625moco) 
summary(mod_4625moco_t1_t2) 
mod_4625moco
mod_4625moco_t1_t2

## Posterior Predictive Checks
pp_check(mod_4625moco)
pp_check(mod_4625moco_t1_t2)

## Generate trace plots
posterior_mod_4625moco <- as.matrix(mod_4625moco)
mcmc_trace(posterior_mod_4625moco, regex_pars = "trt")

posterior_mod_4625moco_t1_t2 <- as.matrix(mod_4625moco_t1_t2)
mcmc_trace(posterior_mod_4625moco_t1_t2, regex_pars = "trt")

## Model performance
(loo_4625moco <- loo(mod_4625moco, k_threshold = 0.7))
(loo_4625moco_t1_t2 <- loo(mod_4625moco_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625moco)
stan_diag(mod_4625moco_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625moco_ar <- mcmc_areas_ridges(posterior_mod_4625moco, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Vehicle vs AZD4625 mono + combo: all studies") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625moco_ar 
ggsave("figures/fit-posterior-azd4625moco-vs-vehicle.png", plt_4625moco_ar)
#
plt_4625moco_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_4625moco, 
    pars = vars(
        starts_with("trt"),
        -c("trtAZD4625:timepoint2weeks:mutationstk11",
           "trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Vehicle vs AZD4625 mono + combo: all studies",
         caption = "Dropped 3-way interactions") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625moco_ar_no3way 
ggsave("figures/fit-posterior-azd4625moco-vs-vehicle-no3way.png",
       plt_4625moco_ar_no3way)
#
plt_4625moco_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625moco_t1_t2, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Vehicle vs AZD4625 mono + combo: baseline and 2 weeks, all studies") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625moco_ar_t1_t2 
ggsave("figures/fit-posterior-azd4625moco-vs-vehicle-t1-t2.png",
       plt_4625moco_ar_t1_t2)
#
plt_4625moco_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_4625moco_t1_t2, 
    pars = vars(
        starts_with("trt"),
        -c("trtAZD4625:timepoint2weeks:mutationstk11",
           "trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Vehicle vs AZD4625 mono + combo: baseline and 2 weeks, all studies",
         caption = "Dropped 3-way interactions") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625moco_ar_t1_t2_no3way
ggsave("figures/fit-posterior-azd4625moco-vs-vehicle-t1-t2-no3way.png",
       plt_4625moco_ar_t1_t2_no3way)
save.image("03-data.RData")
```



### AZD4625 mono vs AZD4625 combo

```{r fit-model-4625-vs-4625combo}
# Extract subset of data
mod_df_4625monoco <- dt %>% 
    filter(str_detect(trt, c("AZD4625"))) %>%
    mutate(trt = relevel(trt, ref = "AZD4625")) %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625monoco <- mod_df_4625monoco %>% distinct(study, mutation)
#
mod_df_4625monoco_t1_t2 <- mod_df_4625monoco %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625monoco_t1_t2 <- mod_df_4625monoco_t1_t2 %>% distinct(study, mutation)

# Fit model
mod_4625monoco <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                           data = mod_df_4625monoco,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)
mod_4625monoco_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                                    (1|mutation/study),
                          data = mod_df_4625monoco_t1_t2,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625monoco) 
summary(mod_4625monoco_t1_t2) 
mod_4625monoco
mod_4625monoco_t1_t2

## Posterior Predictive Checks
pp_check(mod_4625monoco)
pp_check(mod_4625monoco_t1_t2)

## Generate trace plots
posterior_mod_4625monoco <- as.matrix(mod_4625monoco)
mcmc_trace(posterior_mod_4625monoco, regex_pars = "trt")

posterior_mod_4625monoco_t1_t2 <- as.matrix(mod_4625monoco_t1_t2)
mcmc_trace(posterior_mod_4625monoco_t1_t2, regex_pars = "trt")

## Model performance
(loo_4625monoco <- loo(mod_4625monoco, k_threshold = 0.7))
(loo_4625monoco_t1_t2 <- loo(mod_4625monoco_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625monoco)
stan_diag(mod_4625monoco_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625monoco_ar <- mcmc_areas_ridges(posterior_mod_4625monoco, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against AZD4625",
         title = "AZD4625 mono vs AZD4625 combo: all studies") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625monoco_ar 
ggsave("figures/fit-posterior-azd4625combo-vs-azd4625mono.png",
       plt_4625monoco_ar)
#
plt_4625monoco_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_4625monoco, 
    pars = vars(
        starts_with("trt"),
        -c("trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against AZD4625",
         title = "AZD4625 mono vs AZD4625 combo: all studies",
         caption = "Dropped 3-way interactions") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625monoco_ar_no3way 
ggsave("figures/fit-posterior-azd4625combo-vs-azd4625mono-no3way.png",
       plt_4625monoco_ar_no3way)
#
plt_4625monoco_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625monoco_t1_t2, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against AZD4625",
         title = "AZD4625 mono vs AZD4625 combo: baseline and 2 weeks, all studies") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625monoco_ar_t1_t2 
ggsave("figures/fit-posterior-azd4625combo-vs-azd4625mono-t1-t2.png",
       plt_4625monoco_ar_t1_t2)
#
plt_4625monoco_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_4625monoco_t1_t2, 
    pars = vars(
        starts_with("trt"),
        -c("trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against AZD4625",
         title = "AZD4625 mono vs AZD4625 combo: baseline and 2 weeks, all studies",
         caption = "Dropped 3-way interactions") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625monoco_ar_t1_t2_no3way
ggsave("figures/fit-posterior-azd4625combo-vs-azd4625mono-t1-t2-no3way.png",
       plt_4625monoco_ar_t1_t2_no3way)
save.image("03-data.RData")
```



### AZD4625 vs All Other treatments

```{r fit-4625ref-vs-trts}
# Subset data
mod_df_4625ref <- dt %>% 
    mutate(trt = relevel(trt, ref = "AZD4625")) %>% 
    # filter(!timepoint == "3weeks")  %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625ref <- mod_df_4625ref %>% distinct(study, mutation)

mod_df_4625ref_t1_t2 <- mod_df_4625ref %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625ref_t1_t2 <- mod_df_4625ref_t1_t2 %>% distinct(study, mutation)

# Fit model
mod_4625ref <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                               (1|mutation/study),
                           data = mod_df_4625ref,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)
mod_4625ref_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                               (1|mutation/study),
                           data = mod_df_4625ref_t1_t2,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625ref) 
summary(mod_4625ref_t1_t2)
mod_4625ref
mod_4625ref_t1_t2

## Posterior Predictive Checks
pp_check(mod_4625ref)
pp_check(mod_4625ref_t1_t2)

## Model performance
# (loo_4625ref <- loo(mod_4625ref, k_threshold = 0.7))
# (loo_4625ref_t1_t2 <- loo(mod_4625ref_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_4625ref <- as.matrix(mod_4625ref)
mcmc_trace(posterior_mod_4625ref, regex_pars = "trt")

posterior_mod_4625ref_t1_t2 <- as.matrix(mod_4625ref_t1_t2)
mcmc_trace(posterior_mod_4625ref_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_4625ref)
stan_diag(mod_4625ref_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625ref_ar <- mcmc_areas_ridges(posterior_mod_4625ref, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    xlab("Change in log voxels against AZD4625") +
    ggtitle("AZD4625 mono vs All Other Treatments: all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625ref_ar 
ggsave("figures/fit-posterior-trts-vs-azd4625ref.png", plt_4625ref_ar)

plt_4625ref_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_4625ref, 
    pars = vars(starts_with("trt"),
                -c("trtVehicle:timepoint2weeks:mutationstk11",
                   "trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11",
                   "trtCisplatin:timepoint2weeks:mutationstk11")), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against AZD4625",
         title = "AZD4625 mono vs All Other Treatments: all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625ref_ar_no3way 
ggsave("figures/fit-posterior-trts-vs-azd4625ref-no3way.png",
       plt_4625ref_ar_no3way)

plt_4625ref_ar_t1_t2 <- mcmc_areas_ridges(
    posterior_mod_4625ref_t1_t2, 
    pars = vars(starts_with("trt")), 
    prob = 0.95
) +
    xlab("Change in log voxels against AZD4625") +
    ggtitle("Treatments vs AZD4625: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625ref_ar_t1_t2 
ggsave("figures/fit-posterior-trts-vs-azd4625ref-t1-t2.png",
       plt_4625ref_ar_t1_t2)

plt_4625ref_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_4625ref_t1_t2, 
    pars = vars(starts_with("trt"),
                -c("trtVehicle:timepoint2weeks:mutationstk11", 
                   "trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11",
                   "trtCisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against AZD4625",
         title = "Treatments vs AZD4625: baseline and 2 weeks, all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625ref_ar_t1_t2_no3way 
ggsave("figures/fit-posterior-trts-vs-azd4625ref-t1-t2-no3way.png",
       plt_4625ref_ar_t1_t2_no3way)

save.image("03-data.RData")
```


### Vehicle vs AZD6738 mono + combo

```{r fit-model-vehicle-vs-6738_moco}
# Extract subset of data
mod_df_6738moco <- dt %>% 
    filter(str_detect(trt, c("Vehicle|AZD6738"))) %>% 
    # filter(!timepoint == "3weeks")  %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_6738moco <- mod_df_6738moco %>% distinct(study, mutation)

mod_df_6738moco_t1_t2 <- mod_df_6738moco %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_6738moco_t1_t2 <- mod_df_6738moco_t1_t2 %>% distinct(study, mutation)

# Fit model
mod_6738moco <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                   data = mod_df_6738moco,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 6000)
mod_6738moco_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                   data = mod_df_6738moco_t1_t2,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_6738moco)
summary(mod_6738moco_t1_t2)

## Posterior Predictive Checks
pp_check(mod_6738moco)
pp_check(mod_6738moco_t1_t2)

## Model performance
(loo_6738moco_t1_t2 <- loo(mod_6738moco_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_6738moco <- as.matrix(mod_6738moco)
mcmc_trace(posterior_mod_6738moco, regex_pars = "trt")

posterior_mod_6738moco_t1_t2 <- as.matrix(mod_6738moco_t1_t2)
mcmc_trace(posterior_mod_6738moco_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_6738moco)
stan_diag(mod_6738moco_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_6738moco_ar <- mcmc_areas_ridges(posterior_mod_6738moco, 
                                          regex_pars = "trt", 
                                          prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("AZD6738 combo + mono vs Vehicle: all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_6738moco_ar 
ggsave("figures/fit-posterior-azd6738-mono-combo-vs-vehicle.png",
       plt_6738moco_ar)
#
plt_6738moco_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_6738moco_t1_t2, 
                                          regex_pars = "trt", 
                                          prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("AZD6738 combo + mono vs Vehicle: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_6738moco_ar_t1_t2
ggsave("figures/fit-posterior-azd6738-mono-combo-vs-vehicle-t1-t2.png",
       plt_6738moco_ar_t1_t2)

save.image("03-data.RData")
```

 

### Vehicle vs Cisplatin mono + combo

```{r fit-model-vehicle-vs-cisplatin-moco}
# Extract subset of data
mod_df_cis_moco <- dt %>% 
    filter(str_detect(trt, c("Vehicle|Cisplatin"))) %>% 
    # filter(!timepoint == "3weeks")  %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_cis_moco <- mod_df_cis_moco %>% distinct(study, mutation)

mod_df_cis_moco_t1_t2 <- mod_df_cis_moco %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_cis_moco_t1_t2 <- mod_df_cis_moco_t1_t2 %>% distinct(study, mutation)

# Fit model
mod_cis_moco <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                   data = mod_df_cis_moco,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 6000)
mod_cis_moco_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                              (1|mutation/study),
                   data = mod_df_cis_moco_t1_t2,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_cis_moco) 
summary(mod_cis_moco_t1_t2)

## Posterior Predictive Checks
pp_check(mod_cis_moco)
pp_check(mod_cis_moco_t1_t2)

## Model performance
(loo_cis_moco <- loo(mod_cis_moco, k_threshold = 0.7))
(loo_cis_moco_t1_t2 <- loo(mod_cis_moco_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_cis_moco <- as.matrix(mod_cis_moco)
mcmc_trace(posterior_mod_cis_moco, regex_pars = "trt")

posterior_mod_cis_moco_t1_t2 <- as.matrix(mod_cis_moco_t1_t2)
mcmc_trace(posterior_mod_cis_moco_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_cis_moco)
stan_diag(mod_cis_moco_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_cis_moco_ar <- mcmc_areas_ridges(posterior_mod_cis_moco, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("Cisplatin combo + mono vs Vehicle: all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_cis_moco_ar
ggsave("figures/fit-posterior-cisplatin-mono-combo-vs-vehicle.png",
       plt_cis_moco_ar)
#
plt_cis_moco_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_cis_moco, 
    pars = vars(starts_with("trt"),
                -c("trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11",
                   "trtCisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "Cisplatin combo + mono vs Vehicle: all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_cis_moco_ar_no3way
ggsave("figures/fit-posterior-cisplatin-mono-combo-vs-vehicle-no3way.png",
       plt_cis_moco_ar_no3way)
#
plt_cis_moco_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_cis_moco_t1_t2, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("Cisplatin combo + mono vs Vehicle: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_cis_moco_ar_t1_t2
ggsave("figures/fit-posterior-cisplatin-mono-combo-vs-vehicle-t1-t2.png",
       plt_cis_moco_ar_t1_t2)
#
plt_cis_moco_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_cis_moco_t1_t2, 
    pars = vars(starts_with("trt"),
                -c("trtAZD4625 + Cisplatin:timepoint2weeks:mutationstk11",
                   "trtCisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "Cisplatin combo + mono vs Vehicle: baseline and 2 weeks, all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_cis_moco_ar_t1_t2_no3way
ggsave("figures/fit-posterior-cisplatin-mono-combo-vs-vehicle-t1-t2-no3way.png",
       plt_cis_moco_ar_t1_t2_no3way)

save.image("03-data.RData")
```



### Vehicle vs AZD4625, Cisplatin, Cisplatin + PD-L1 combo

```{r fit_model_veh_4625_cisplatin-moco}
# Extract subset of data
mod_df_4625cis_moco <- dt %>% 
    filter(
        trt %in% c("Vehicle", "AZD4625", "Cisplatin", "Cisplatin + PD-L1")
    ) %>%
    # filter(!timepoint == "3weeks")  %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625cis_moco <- mod_df_4625cis_moco %>% distinct(study, mutation)

mod_df_4625cis_moco_t1_t2 <- mod_df_4625cis_moco %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625cis_moco_t1_t2 <- mod_df_4625cis_moco_t1_t2 %>% 
    distinct(study, mutation)

# Fit model
mod_4625cis_moco <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                                  (1|mutation/study),
                              data = mod_df_4625cis_moco,
                              adapt_delta = 0.9999,
                              seed = 1234,
                              iter = 6000)
mod_4625cis_moco_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint * mutation +
                                  (1|mutation/study),
                              data = mod_df_4625cis_moco_t1_t2,
                              adapt_delta = 0.9999,
                              seed = 1234,
                              iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625cis_moco) 
summary(mod_4625cis_moco_t1_t2) 

## Posterior Predictive Checks
pp_check(mod_4625cis_moco)
pp_check(mod_4625cis_moco_t1_t2)

## Model performance
(loo_4625cis_moco <- loo(mod_4625cis_moco, k_threshold = 0.7))
(loo_4625cis_moco_t1_t2 <- loo(mod_4625cis_moco_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_4625cis_moco <- as.matrix(mod_4625cis_moco)
mcmc_trace(posterior_mod_4625cis_moco, regex_pars = "trt")

posterior_mod_4625cis_moco_t1_t2 <- as.matrix(mod_4625cis_moco_t1_t2)
mcmc_trace(posterior_mod_4625cis_moco_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_4625cis_moco)
stan_diag(mod_4625cis_moco_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625cis_moco_ar <- mcmc_areas_ridges(posterior_mod_4625cis_moco, 
                                              regex_pars = "trt", 
                                              prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("AZD4625, Cisplatin, Cisplatin + PD-L1 vs Vehicle: all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625cis_moco_ar 
ggsave("figures/fit-posterior-4625cisplatin-moco-vs-vehicle.png",
       plt_4625cis_moco_ar)
#
plt_4625cis_moco_ar_no3way <- mcmc_areas_ridges(
    posterior_mod_4625cis_moco, 
    pars = vars(starts_with("trt"), 
                -c("trtAZD4625:timepoint2weeks:mutationstk11",
                   "trtCisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625, Cisplatin, Cisplatin + PD-L1 vs Vehicle: all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625cis_moco_ar_no3way 
ggsave("figures/fit-posterior-4625cisplatin-moco-vs-vehicle-no3way.png",
       plt_4625cis_moco_ar_no3way)
#
plt_4625cis_moco_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625cis_moco_t1_t2, 
                                              regex_pars = "trt", 
                                              prob = 0.95) +
    xlab("Change in log voxels against Vehicle") +
    ggtitle("AZD4625, Cisplatin, Cisplatin + PD-L1 vs Vehicle: baseline and 2 weeks, all studies") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625cis_moco_ar_t1_t2 
ggsave("figures/fit-posterior-4625cisplatin-moco-vs-vehicle-t1-t2.png",
       plt_4625cis_moco_ar_t1_t2)
#
plt_4625cis_moco_ar_t1_t2_no3way <- mcmc_areas_ridges(
    posterior_mod_4625cis_moco_t1_t2, 
    pars = vars(starts_with("trt"),
                -c("trtAZD4625:timepoint2weeks:mutationstk11",
                   "trtCisplatin:timepoint2weeks:mutationstk11")
    ), 
    prob = 0.95
) +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD4625, Cisplatin, Cisplatin + PD-L1 vs Vehicle: baseline and 2 weeks, all studies",
         caption = "Dropped 3-way interactions") +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625cis_moco_ar_t1_t2_no3way 
ggsave("figures/fit-posterior-4625cisplatin-moco-vs-vehicle-t1-t2-no3way.png",
       plt_4625cis_moco_ar_t1_t2_no3way)

save.image("03-data.RData")
```


## Conclusions for all fitted models above

All MCSEs were 0.0 except for log-posterior, which is 0.1. Thus, no sample had substantial level of noise. 

Rhat values were all 1.0, indicating convergence and equilibrium of
simulated chains. 

The size of the ESSs were large, suggesting the simulated
chains were well mixed. 

The trace plots confirm convergence and mixing.   

The posterior predictive plots shows fairly close agreement between the model 
and data. The results of the `loo(.)` suggests that the predictive performance
of the fitted models were well assessed.   

The diagnostic plots indicates fairly explored posterior distribution.   



# Session info

```{r ws-info, echo=FALSE, include = FALSE }
save.image("03-data.RData")

sessionInfo()
```
