---
title: "Historic data in vehicle + Alpha-treated GEMMs"
subtitle: "Statistical analysis of data from Sv"
author: "Gabriel Teku"
date: "2024-03-06"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(tidyverse)
library(emmeans)
library(rstan)
library(rstanarm)
library(bayesplot)
library(brms)
library(loo)

set.seed <- 1234

## colourblind-friendly colours for ggplot
cbb_palette <- c(
  "#000000", "#E69F00", "#56B4E9", "#F0E442", "#0072B2",
  "#D55E00", "#CC79A7", "#0099C6", "#C7A92A", "#009E73",
  "#5A9F9F", "#F8766D", "#C49A00", "#F2C80F", "#993D3D",
  "#4F5D75", "#0089A3", "#A8A878", "#63C5DA", "#79A779"
)

options(digits = 5, tibble.print_max = 10^6, mc.cores = 4)

if (file.exists("05-data.RData")) load("05-data.RData")
```


# Load and filter data

```{r load-cleaned-data, echo=FALSE}
dt <- read_rds("clean-data-se.rds")
```


# Statistical inference 

## Fit Bayesian mixed effects model 

Interested in picking up trt effects & effects of treatment over time. 
Does this differ between mutation statuses?
Consider study as random effect with sampling from the entire study space.

### Vehicle vs Alpha mono

```{r fit-vehicle-vs-Alpha-mono_se}
# Extract subset of data
mod_df_4625mono_se <- dt %>% 
    filter(trt %in% c("Vehicle", "Alpha")) %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove)) %>% 
    droplevels()
studies_df_4625mono_se <- mod_df_4625mono_se %>% distinct(study)
studies_df_4625mono_se

mod_df_4625mono_se_t1_t2 <- mod_df_4625mono_se %>% 
    filter(timepoint %in% c("baseline", "1week") ) %>% droplevels()
studies_df_4625mono_se_t1_t2 <- mod_df_4625mono_se_t1_t2 %>% distinct(study)
studies_df_4625mono_se_t1_t2

# Fit model
mod_4625mono_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                          data = mod_df_4625mono_se,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 5000)
mod_4625mono_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                          data = mod_df_4625mono_se_t1_t2,
                          adapt_delta = 0.999,
                          seed = 1234,
                          iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625mono_se)
summary(mod_4625mono_se_t1_t2)
mod_4625mono_se
mod_4625mono_se_t1_t2

## Posterior Predictive Checks
pp_check(mod_4625mono_se)
pp_check(mod_4625mono_se_t1_t2)

## Generate trace plots
posterior_mod_4625mono_se <- as.matrix(mod_4625mono_se)
mcmc_trace(posterior_mod_4625mono_se, regex_pars = "trt")

posterior_mod_4625mono_se_t1_t2 <- as.matrix(mod_4625mono_se_t1_t2)
mcmc_trace(posterior_mod_4625mono_se_t1_t2, regex_pars = "trt")

## Model performance
# (loo_4625mono_se <- loo(mod_4625mono_se, k_threshold = 0.7))
# (loo_4625mono_se_t1_t2 <- loo(mod_4625mono_se_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625mono_se)
stan_diag(mod_4625mono_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625mono_se_ar <- mcmc_areas_ridges(posterior_mod_4625mono_se, 
                       regex_pars = "trt",
                       prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Alpha mono vs Vehicle: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625mono_se_ar
ggsave("figures-se-se/fit-posterior-Alpha-vehicle-se.png", plt_4625mono_se_ar)
#
plt_4625mono_se_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625mono_se_t1_t2, 
                       regex_pars = "trt", 
                       prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Alpha mono vs Vehicle: baseline and 1 week, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625mono_se_ar_t1_t2
ggsave("figures-se/fit-posterior-Alpha-vehicle-se.png", plt_4625mono_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```


### Vehicle vs Alpha combo

```{r fit-model-vehicle-vs-4625combo_se}
# Extract subset of data
mod_df_4625combo_se <- dt %>% 
    filter(str_detect(trt, c("Vehicle|Alpha +"))) %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625combo_se <- mod_df_4625combo_se %>% distinct(study)
#
mod_df_4625combo_se_t1_t2 <- mod_df_4625combo_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625combo_se_t1_t2 <- mod_df_4625combo_se_t1_t2 %>% distinct(study)

# Fit model
mod_4625combo_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                           data = mod_df_4625combo_se,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 6000)
mod_4625combo_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                          data = mod_df_4625combo_se_t1_t2,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625combo_se) 
summary(mod_4625combo_se_t1_t2) 

## Posterior Predictive Checks
pp_check(mod_4625combo_se)
pp_check(mod_4625combo_se_t1_t2)

## Generate trace plots
posterior_mod_4625combo_se <- as.matrix(mod_4625combo_se)
mcmc_trace(posterior_mod_4625combo_se, regex_pars = "trt")

posterior_mod_4625combo_se_t1_t2 <- as.matrix(mod_4625combo_se_t1_t2)
mcmc_trace(posterior_mod_4625combo_se_t1_t2, regex_pars = "trt")

## Model performance
# (loo_4625combo_se <- loo(mod_4625combo_se, k_threshold = 0.7))
# (loo_4625combo_se_t1_t2 <- loo(mod_4625combo_se_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625combo_se)
stan_diag(mod_4625combo_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625combo_se_ar <- mcmc_areas_ridges(posterior_mod_4625combo_se, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Alpha combo vs Vehicle: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625combo_se_ar 
ggsave("figures-se/fit-posterior-Alphacombo_se-vs-vehicle-se.png", plt_4625combo_se_ar)

plt_4625combo_se_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625combo_se_t1_t2, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Alpha combo vs Vehicle: baseline and 2 weeks, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625combo_se_ar_t1_t2 
ggsave("figures-se/fit-posterior-Alphacombo_se-vs-vehicle-t1-t2-se.png",
       plt_4625combo_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```


### Vehicle vs Alpha mono + combo

```{r fit-model-vehicle-vs-4625moco_se}
# Extract subset of data
mod_df_4625moco_se <- dt %>% 
    filter(str_detect(trt, c("Vehicle|Alpha"))) %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625moco_se <- mod_df_4625moco_se %>% distinct(study)
#
mod_df_4625moco_se_t1_t2 <- mod_df_4625moco_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625moco_se_t1_t2 <- mod_df_4625moco_se_t1_t2 %>% distinct(study)

# Fit model
mod_4625moco_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                           data = mod_df_4625moco_se,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)
mod_4625moco_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                          data = mod_df_4625moco_se_t1_t2,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625moco_se) 
summary(mod_4625moco_se_t1_t2) 

## Posterior Predictive Checks
pp_check(mod_4625moco_se)
pp_check(mod_4625moco_se_t1_t2)

## Generate trace plots
posterior_mod_4625moco_se <- as.matrix(mod_4625moco_se)
mcmc_trace(posterior_mod_4625moco_se, regex_pars = "trt")

posterior_mod_4625moco_se_t1_t2 <- as.matrix(mod_4625moco_se_t1_t2)
mcmc_trace(posterior_mod_4625moco_se_t1_t2, regex_pars = "trt")

## Model performance
# (loo_4625moco_se <- loo(mod_4625moco_se, k_threshold = 0.7))
# (loo_4625moco_se_t1_t2 <- loo(mod_4625moco_se_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625moco_se)
stan_diag(mod_4625moco_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625moco_se_ar <- mcmc_areas_ridges(posterior_mod_4625moco_se, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Vehicle vs Alpha mono + combo: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625moco_se_ar 
ggsave("figures-se/fit-posterior-Alphamoco_se-vs-vehicle-se.png", plt_4625moco_se_ar)

plt_4625moco_se_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_4625moco_se_t1_t2, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Vehicle vs Alpha mono + combo: baseline and 2 weeks, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625moco_se_ar_t1_t2 
ggsave("figures-se/fit-posterior-Alphamoco_se-vs-vehicle-t1-t2-se.png",
       plt_4625moco_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```




### Alpha mono vs Alpha combo

```{r fit-model-4625mono-vs-4625combo_se}
# Extract subset of data
mod_df_4625monoco_se <- dt %>% 
    filter(str_detect(trt, c("Alpha"))) %>%
    mutate(trt = relevel(trt, ref = "Alpha")) %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625monoco_se <- mod_df_4625monoco_se %>% distinct(study)
#
mod_df_4625monoco_se_t1_t2 <- mod_df_4625monoco_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625monoco_se_t1_t2 <- mod_df_4625monoco_se_t1_t2 %>% distinct(study)

# Fit model
mod_4625monoco_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                           data = mod_df_4625monoco_se,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)
mod_4625monoco_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                          data = mod_df_4625monoco_se_t1_t2,
                          adapt_delta = 0.9999,
                          seed = 1234,
                          iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625monoco_se) 
summary(mod_4625monoco_se_t1_t2) 

## Posterior Predictive Checks
pp_check(mod_4625monoco_se)
pp_check(mod_4625monoco_se_t1_t2)

## Generate trace plots
posterior_mod_4625monoco_se <- as.matrix(mod_4625monoco_se)
mcmc_trace(posterior_mod_4625monoco_se, regex_pars = "trt")

posterior_mod_4625monoco_se_t1_t2 <- as.matrix(mod_4625monoco_se_t1_t2)
mcmc_trace(posterior_mod_4625monoco_se_t1_t2, regex_pars = "trt")

## Model performance
# (loo_4625monoco_se <- loo(mod_4625monoco_se, k_threshold = 0.7))
# (loo_4625monoco_se_t1_t2 <- loo(mod_4625monoco_se_t1_t2, k_threshold = 0.7))

## More diagnostics
stan_diag(mod_4625monoco_se)
stan_diag(mod_4625monoco_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625monoco_se_ar <- mcmc_areas_ridges(posterior_mod_4625monoco_se, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Alpha mono",
         title = "Alpha mono vs Alpha combo: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625monoco_se_ar 
ggsave("figures-se/fit-posterior-Alphacombo-vs-Alphamono-se.png", 
       plt_4625monoco_se_ar)

plt_4625monoco_se_ar_t1_t2 <- mcmc_areas_ridges(
    posterior_mod_4625monoco_se_t1_t2, 
    regex_pars = "trt", 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Alpha mono",
         title = "Alpha mono vs Alpha combo: baseline and 2 weeks, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625monoco_se_ar_t1_t2 
ggsave("figures-se/fit-posterior-Alphacombo-vs-Alphamono-t1-t2-se.png",
       plt_4625monoco_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```



### Alpha mono vs All Other treatments

```{r fit-4625ref_se-vs-trts}
# Subset data
mod_df_4625ref_se <- dt %>% 
    mutate(trt = relevel(trt, "Alpha")) %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625ref_se <- mod_df_4625ref_se %>% distinct(study, mutation)

mod_df_4625ref_se_t1_t2 <- mod_df_4625ref_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625ref_se_t1_t2 <- mod_df_4625ref_se_t1_t2 %>% 
    distinct(study, mutation)

# Fit model
mod_4625ref_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                           data = mod_df_4625ref_se,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)
mod_4625ref_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                           data = mod_df_4625ref_se_t1_t2,
                           adapt_delta = 0.9999,
                           seed = 1234,
                           iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625ref_se)
summary(mod_4625ref_se_t1_t2)

## Posterior Predictive Checks
pp_check(mod_4625ref_se)
pp_check(mod_4625ref_se_t1_t2)

## Model performance
# (loo_4625ref_se <- loo(mod_4625ref_se, k_threshold = 0.7))
# (loo_4625ref_se_t1_t2 <- loo(mod_4625ref_se_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_4625ref_se <- as.matrix(mod_4625ref_se)
mcmc_trace(posterior_mod_4625ref_se, regex_pars = "trt")

posterior_mod_4625ref_se_t1_t2 <- as.matrix(mod_4625ref_se_t1_t2)
mcmc_trace(posterior_mod_4625ref_se_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_4625ref_se)
stan_diag(mod_4625ref_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625ref_se_ar <- mcmc_areas_ridges(posterior_mod_4625ref_se, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Alpha: Sweden studies",
         title = "Alpha mono vs All Other treatments: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625ref_se_ar 
ggsave("figures-se/fit-posterior-trts-vs-Alpharef_se.png", plt_4625ref_se_ar)
#
plt_4625ref_se_ar_t1_t2 <- mcmc_areas_ridges(
    posterior_mod_4625ref_se_t1_t2, 
    pars = vars(starts_with("trt")), 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Alpha",
         title = "Alpha mono vs All Other Treatments: baseline and 2 weeks, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625ref_se_ar_t1_t2 
ggsave("figures-se/fit-posterior-trts-vs-Alpharef_se-t1-t2.png",
       plt_4625ref_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```


### Vehicle vs AZD6738 mono + combo

```{r fit-model-vehicle-vs-6738_moco-se}
# Extract subset of data
mod_df_6738moco_se <- dt %>% 
    filter(str_detect(trt, c("Vehicle|AZD6738"))) %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_6738moco_se <- mod_df_6738moco_se %>% distinct(study)

mod_df_6738moco_se_t1_t2 <- mod_df_6738moco_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_6738moco_se_t1_t2 <- mod_df_6738moco_se_t1_t2 %>% distinct(study)

# Fit model
mod_6738moco_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                   data = mod_df_6738moco_se,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 5000)
mod_6738moco_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                   data = mod_df_6738moco_se_t1_t2,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_6738moco_se)
summary(mod_6738moco_se_t1_t2)

## Posterior Predictive Checks
pp_check(mod_6738moco_se)
pp_check(mod_6738moco_se_t1_t2)

## Model performance
# (loo_6738moco_se <- loo(mod_6738moco_se, k_threshold = 0.7))
# (loo_6738moco_se_t1_t2 <- loo(mod_6738moco_se_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_6738moco_se <- as.matrix(mod_6738moco_se)
mcmc_trace(posterior_mod_6738moco_se, regex_pars = "trt")

posterior_mod_6738moco_se_t1_t2 <- as.matrix(mod_6738moco_se_t1_t2)
mcmc_trace(posterior_mod_6738moco_se_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_6738moco_se)
stan_diag(mod_6738moco_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_6738moco_se_ar <- mcmc_areas_ridges(posterior_mod_6738moco_se, 
                                          regex_pars = "trt", 
                                          prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD6738 combo + mono vs Vehicle: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_6738moco_se_ar 
ggsave("figures-se/fit-posterior-azd6738-mono-combo-vs-vehicle-se.png",
       plt_6738moco_se_ar)
#
plt_6738moco_se_ar_t1_t2 <- mcmc_areas_ridges(posterior_mod_6738moco_se_t1_t2, 
                                          regex_pars = "trt", 
                                          prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "AZD6738 combo + mono vs Vehicle: baseline and 2 weeks, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_6738moco_se_ar_t1_t2
ggsave("figures-se/fit-posterior-azd6738-mono-combo-vs-vehicle-t1-t2-se.png",
       plt_6738moco_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```

 

### Vehicle vs Cisplatin combo

```{r fit-model-vehicle-vs-cisplatin-combo-se}
# Extract subset of data
mod_df_cis_combo_se <- dt %>% 
    filter(str_detect(trt, c("Vehicle|Cisplatin"))) %>% 
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_cis_combo_se <- mod_df_cis_combo_se %>% distinct(study)

mod_df_cis_combo_se_t1_t2 <- mod_df_cis_combo_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_cis_combo_se_t1_t2 <- mod_df_cis_combo_se_t1_t2 %>% distinct(study)

# Fit model
mod_cis_combo_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                   data = mod_df_cis_combo_se,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 6000)
mod_cis_combo_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                   data = mod_df_cis_combo_se_t1_t2,
                   adapt_delta = 0.9999,
                  seed = 1234,
                   iter = 6000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_cis_combo_se) 
summary(mod_cis_combo_se_t1_t2)

## Posterior Predictive Checks
pp_check(mod_cis_combo_se)
pp_check(mod_cis_combo_se_t1_t2)

## Model performance
# (loo_cis_combo_se <- loo(mod_cis_combo_se, k_threshold = 0.7))
# (loo_cis_combo_se_t1_t2 <- loo(mod_cis_combo_se_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_cis_combo_se <- as.matrix(mod_cis_combo_se)
mcmc_trace(posterior_mod_cis_combo_se, regex_pars = "trt")

posterior_mod_cis_combo_se_t1_t2 <- as.matrix(mod_cis_combo_se_t1_t2)
mcmc_trace(posterior_mod_cis_combo_se_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_cis_combo_se)
stan_diag(mod_cis_combo_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_cis_combo_se_ar <- mcmc_areas_ridges(posterior_mod_cis_combo_se, 
                                  regex_pars = "trt", 
                                  prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Cisplatin combo vs Vehicle: Sweden studies", 
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_cis_combo_se_ar
ggsave("figures-se/fit-posterior-cisplatin-mono-combo-vs-vehicle-se.png",
       plt_cis_combo_se_ar)
#
plt_cis_combo_se_ar_t1_t2 <- mcmc_areas_ridges(
    posterior_mod_cis_combo_se_t1_t2, 
    regex_pars = "trt", 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") + 
    labs(x = "Change in log voxels against Vehicle",
         title = "Cisplatin combo vs Vehicle: baseline and 2 weeks, Sweden studies", 
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_cis_combo_se_ar_t1_t2
ggsave("figures-se/fit-posterior-cisplatin-mono-combo-vs-vehicle-t1-t2-se.png",
       plt_cis_combo_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```



### Vehicle vs Alpha, Cisplatin, Cisplatin + PD-L1 combo

```{r fit_model_veh_4625_cisplatin-moco-se}
# Extract subset of data
mod_df_4625cis_moco_se <- dt %>% 
    filter(
        trt %in% c("Vehicle", "Alpha", "Cisplatin + PD-L1")
    ) %>%
    group_by(trt, timepoint, mutation) %>% 
    mutate(perc_na = sum(is.na(log_voxels)) / n(),
           na_remove = perc_na > 0.5) %>% 
    filter(!na_remove) %>% 
    filter(n() >= 3) %>% 
    ungroup() %>% 
    select(-c(perc_na, na_remove))
studies_df_4625cis_moco_se <- mod_df_4625cis_moco_se %>% distinct(study)

mod_df_4625cis_moco_se_t1_t2 <- mod_df_4625cis_moco_se %>% 
    filter(timepoint %in% c("baseline", "2weeks") )
studies_df_4625cis_moco_se_t1_t2 <- mod_df_4625cis_moco_se_t1_t2 %>%
    distinct(study)

# Fit model
mod_4625cis_moco_se <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                              data = mod_df_4625cis_moco_se,
                              adapt_delta = 0.9999,
                              seed = 1234,
                              iter = 5000)
mod_4625cis_moco_se_t1_t2 <- stan_lmer(log_voxels ~ trt * timepoint + (1|study),
                              data = mod_df_4625cis_moco_se_t1_t2,
                              adapt_delta = 0.9999,
                              seed = 1234,
                              iter = 5000)

# Diagnostics ---------------------------------------------------------------
## Evaluate convergence
summary(mod_4625cis_moco_se) 
summary(mod_4625cis_moco_se_t1_t2) 

## Posterior Predictive Checks
pp_check(mod_4625cis_moco_se)
pp_check(mod_4625cis_moco_se_t1_t2)

## Model performance
# (loo_4625cis_moco_se <- loo(mod_4625cis_moco_se, k_threshold = 0.7))
# (loo_4625cis_moco_se_t1_t2 <- loo(mod_4625cis_moco_se_t1_t2, k_threshold = 0.7))

## Generate trace plots
posterior_mod_4625cis_moco_se <- as.matrix(mod_4625cis_moco_se)
mcmc_trace(posterior_mod_4625cis_moco_se, regex_pars = "trt")

posterior_mod_4625cis_moco_se_t1_t2 <- as.matrix(mod_4625cis_moco_se_t1_t2)
mcmc_trace(posterior_mod_4625cis_moco_se_t1_t2, regex_pars = "trt")

## More diagnostics
stan_diag(mod_4625cis_moco_se)
stan_diag(mod_4625cis_moco_se_t1_t2)

# Extract posterior samples -------------------------------------------------
plt_4625cis_moco_se_ar <- mcmc_areas_ridges(posterior_mod_4625cis_moco_se, 
                                              regex_pars = "trt", 
                                              prob = 0.95) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Alpha, Cisplatin + PD-L1 vs Vehicle: Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625cis_moco_se_ar 
ggsave("figures-se/fit-posterior-4625cisplatin-moco-vs-vehicle-se.png",
       plt_4625cis_moco_se_ar)
#
plt_4625cis_moco_se_ar_t1_t2 <- mcmc_areas_ridges(
    posterior_mod_4625cis_moco_se_t1_t2, 
    regex_pars = "trt", 
    prob = 0.95
) +
    geom_vline(aes(xintercept = 0), linetype = "dashed", color = "red") +
    labs(x = "Change in log voxels against Vehicle",
         title = "Alpha, Cisplatin + PD-L1 vs Vehicle: baseline and 2 weeks, Sweden studies",
         caption = "Studies performed in Sweden") +
    theme_bw() +
    theme(text = element_text(face = "bold"))
plt_4625cis_moco_se_ar_t1_t2 
ggsave("figures-se/fit-posterior-4625cisplatin-moco-vs-vehicle-t1-t2-se.png",
       plt_4625cis_moco_se_ar_t1_t2)

save.image("05-data.RData")
beepr::beep()
```


## Conclusions for all fitted models above

All MCSEs were 0.0 except for log-posterior, which is 0.1. Thus, no sample had substantial level of noise. 

Rhat values were all 1.0, indicating convergence and equilibrium of
simulated chains. 

The size of the ESSs were large, suggesting the simulated
chains were well mixed. 

The trace plots confirm convergence and mixing.   

The posterior predictive plots shows fairly close agreement between the model 
and data. The results of the `loo(.)` suggests that the predictive performance
of the fitted models were well assessed.   

The diagnostic plots indicates fairly explored posterior distribution.   



# Session info

```{r ws-info, echo=FALSE, include = FALSE }
save.image("05-data.RData")

sessionInfo()
```
